

# Sentiment analysis with tidy data  

```{r}
library(tidytext)
library(tidyverse)  
```

As discussed above, there are a variety of methods and dictionaries that exist for evaluating the opinion or emotion in text. The tidytext package contains several sentiment lexicons. Three general-purpose lexicons are

* `AFINN` from [Finn Årup Nielsen](http://www2.imm.dtu.dk/pubdb/views/publication_details.php?id=6010),
* `bing` from [Bing Liu and collaborators](https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html), and
* `nrc` from [Saif Mohammad and Peter Turney](http://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm).  
* `loughran`  


```{r}
get_sentiments("bing")
```

```{r}
# install.packages("textdata")
get_sentiments("afinn")
```

```{r}
get_sentiments("nrc")
```


```{r}
get_sentiments("loughran")
```


## Sentiment analysis with inner join  

```{r}
library(janeaustenr)  

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

```


```{r}
tidy_books
```


```{r}
(nrc_joy <- get_sentiments("nrc") %>%
  filter(sentiment == "joy"))
```

```{r}
tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```



```{r}
tidy_books %>%
  inner_join(get_sentiments("bing")) %>% 
  mutate(index = linenumber %/% 80) %>% 
  count(book, index, sentiment)
```


```{r}
(jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative))
```


```{r,fig.height = 10}
ggplot(jane_austen_sentiment, aes(index, sentiment)) + 
  geom_col(aes(fill = book), show.legend = F) + 
  facet_wrap( ~ book, ncol = 2, scales = "free_x") + 
  theme_classic() 
```






## Most common positive and negative words 

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()
```


```{r}
bing_word_counts
```

我们可以看到情感分析在这里犯的一个小错误，“miss” 这个词被视为负面的，但该词在 Jane Austen 中实际上是对年轻、未婚女士的称呼。我们可以用 `bind_words()` 把 miss 添加到停用词 `stop_words` 中：  

```{r}
custom_stop_words <- tibble(word = c("miss"), lexicon = c("custom")) %>% 
  bind_rows(stop_words)

bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>% 
  anti_join(custom_stop_words) %>%
  group_by(sentiment) %>%
  count(word, sentiment, sort = T) %>% 
  ungroup()
```

```{r}
bing_word_counts %>%
  group_by(sentiment) %>% 
  top_n(10, wt = n) %>%
  ggplot(aes(fct_reorder(word, n), n)) + 
  geom_col(aes(fill = sentiment), show.legend = F) + 
  scale_fill_manual(values = c("#377EB8", "#E41A1C")) + 
  coord_flip() + 
  facet_wrap(~ sentiment, scales = "free_y") + 
  labs(x = NULL, y = NULL,
       title = "Top 10 words of sentiment in Jane Austen's books") + 
  theme_classic() + 
  theme(axis.text.y = element_text(face = "bold.italic", size = 12))
```


## 词云  


```{r}
library(wordcloud2)
tidy_books %>% 
  anti_join(stop_words) %>% 
  inner_join(get_sentiments("bing")) %>% 
  count(sentiment, word, sort = T) %>% 
  .[c(2, 3)] %>% 
  wordcloud2()
```






