

# (PART) Gallery {-}

# Sentiment analysis with `ggpage`

```{r}
knitr::opts_chunk$set(cache = TRUE)
```


```{r}
library(ggpage)
```


```{r}
tinderboxtinderbox
```


`ggpage_quick()` gives a quick default plot like `ggplot::qplot()`, each row of `text` is represented as each row in the plot, the length of each rectangle correponding to the length of each word  

```{r}
ggpage_quick(head(tinderbox))
```


When the data is long enough, `text` is splitting into multiple box of "pages":  

```{r}
ggpage_quick(head(tinderbox, 100))
```

The same result would be achieved by using

```r
tinderbox %>% 
  ggpage_build() %>% 
  ggpage_plot()
``` 

But the second way allows for control over how the text data can be visualized, as we can introduce more code between `ggpage_build` and `ggpage_plot`  

`ggbuld_build()` does the preprocessing job. The first argument can be either a character or a data frame, it can either have each element be a separate line or having each element being separate words. If it is the former case, then `ggbuild_build` split the text column into word tokens. Then it  determines the coordinates of each word

```{r}
tinderbox %>%
  ggpage_build()
```

Map word length to the color of the corresponding rectangle 

```{r, fig.height = 10, fig.length = 8}
tinderbox %>%
  ggpage_build() %>%
  mutate(long_word = str_length(word) > 8) %>%
  ggpage_plot(aes(fill = long_word)) +
  labs(title = "Longer words throughout The Tinder-box") + 
  scale_fill_manual(values = c("grey70", "blue"),
                    labels = c("8 or less", "9 or more"),
                    name = "Word length") + 
  theme(plot.title = element_text(size = 30))
```



```{r}
library(paletteer)
sentiment_types <- get_sentiments("nrc") %>%
  pull(sentiment) %>%
  unique()

prebuild <- imap_dfr(sentiment_types,
  ~ ggpage_build(tinderbox) %>%
    left_join(filter(get_sentiments("nrc"), sentiment == .x), by = "word") %>%
    mutate(sentiment_types = .x,
           score = as.numeric(!is.na(sentiment)),
           score_smooth =slider::slide_dbl(score, mean, .before = 2, .after = 2)))

prebuild
```

```{r, fig.height = 10, fig.width = 10}
library(gganimate)

prebuild %>% 
  ggpage_plot(aes(fill = score_smooth), page.number = "top-left") +
  scale_fill_paletteer_c("viridis::cividis", direction = -1) +
  guides(fill = "none") +
  transition_states(
    sentiment_types,
    transition_length = 10,
    state_length = 5
    ) +
  labs(title = "Sections with a sentiment of {closest_state}\nIn H.C. Andersen's Tinderbox")
```

