

# Text processing examples in R 

 A manual of problem-solving techniques in some common use cases dealing with text data, employing tools like [stringr](https://stringr.tidyverse.org/index.html) and [textclean](https://github.com/trinker/textclean)


## Replacing and removing  

## Combining and splitting 

Combine multiple line tokens to one paragraph (with a defined length) and the other way around  

```{r}
library(ggpage)

tinderbox_line <- head(tinderbox)
tinderbox_paragraph <- head(tinderbox_paragraph) 
  
  
# combine all six rows into a paragraph
tinderbox_line %>% 
  summarize(paragraph = str_c(text, collapse = " "))

# split one paragraph into multiple rows 
# ggpage::nest_paragraphs() extends str_wrap()
tinderbox_paragraph %>% 
  mutate(paragraph_length = str_length(text)) %>% 
  nest_paragraphs(text, width = 50)  # no more than 50 characters per row
```

## Extracting text from pdf and other files  

```{r}
library(pdftools)
```

```{r, eval = FALSE}
download.file("http://arxiv.org/pdf/1403.2805.pdf", "data/1403.2805.pdf", mode = "wb")
```


```{r}
txt <- pdf_text("data/1403.2805.pdf")

# all 29 pages  
length(txt) 

cat(txt[[1]])

enframe(txt) %>% 
  rename(page = name) %>% 
  nest_paragraphs(input = value, width = 100) %>%
  head()
```

### Office documents  


### Images